---
title: "Worksheet-4a in R"
author: "Erl Syron R. Espadon"
output: pdf_document
---



#1. The table below shows the data about shoe size and height. Create a data frame.
#a.	Describe the data.
```{r}
# Create a data frame from the table
shoes <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)

# Print the data frame
print(shoes)
```

#b.	Create a subset by males and females with their corresponding shoe size and height.What its result? Show the R scripts.
```{r}
# Create subsets for males and females
males <- shoes[shoes$Gender == "M", c("Shoe_size", "Height")]
females <- shoes[shoes$Gender == "F", c("Shoe_size", "Height")]

# Print the subsets
print(males)
print(females)
```

#c.	Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
# Calculate the mean of shoe size and height
mean_shoe_size <- mean(shoes$Shoe_size)
mean_height <- mean(shoes$Height)

# Print the means
print(paste("Mean shoe size:", mean_shoe_size))
print(paste("Mean height:", mean_height))
```

#d.	Is there a relationship between shoe size and height? Why?
```{r}
#the taller, the bigger the shoe size.
```



#2.	Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.
```{r}
months <- c("March", "April", "January", "November", "January",
            "September", "October", "September", "November", "August",
            "January", "November", "November", "February", "May", "August",
            "July", "December", "August", "August", "September", "November", 
            "February", "April")
factor_months_vector <- factor(months)
print(factor_months_vector)
```


#3.	Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months)
summary(factor_months_vector)

#In this case, factor_months_vector is more useful than months_vector. The summary of the factor vector provides valuable information about the distribution of the months, while the summary of the character vector only provides basic information about the vector itself.
```


#4.	Create a vector and factor for the table below.
```{r}
direction_vector <- c("East", "West", "West", "West", "West", "North", "North", "North")
factor_data <- factor(direction_vector)
new_order_data <- factor(factor_data, levels = c("East", "West", "North"))
print(new_order_data)
```


#5
#a. a.	Import the excel file into the Environment Pane using read.table() function.Write the code
```{r}
march_data <- read.table("import_march.csv", header = TRUE, sep = ",")
```
#b. View the dataset. Write the R scripts and its result.
```{r}
print(march_data)
```



#6. Full Search (Exhaustive Search Function)
# Get user input
```{r}
exhaustive_search <- function(number) {
  if (number < 1 || number > 50) {
    return("The number selected is beyond the range of 1 to 50")
  } else if (number == 20) {
    return(TRUE)
  } else {
    return(number)
  }
}

# Get user input
user_number <- readline("Enter a number between 1 and 50: ")

# Check if the input is a valid number
if (is.numeric(user_number)) {
  user_number <- as.numeric(user_number)
  # Call the function and display the result
  result <- exhaustive_search(user_number)
  print(result)
} else {
  print("Invalid input. Please enter a number.")
}
```


#7. Change (Minimum Bills Function)
```{r}
min_bills <- function(price) {
  bill_values <- c(1000, 500, 200, 100, 50)
  bills_needed <- rep(0, length(bill_values))
  
  for (i in 1:length(bill_values)) {
    while (price >= bill_values[i]) {
      bills_needed[i] <- bills_needed[i] + 1
      price <- price - bill_values[i]
    }
  }
  
  return(sum(bills_needed))
}

# Get a random price divisible by 50
price <- sample(seq(50, 1000, by = 50), 1)

# Call the function and display the result
num_bills <- min_bills(price)
print(paste("Minimum bills needed:", num_bills))
```


#8.
#a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
# Create the dataframe
grades <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

# Print the dataframe
print(grades)
```
#b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output.
```{r}
for (i in 1:nrow(grades)) {
  student_name <- grades$Name[i]
  average_score <- (grades$Grade1[i] + grades$Grade2[i] + grades$Grade3[i] + grades$Grade4[i]) / 4
  if (average_score > 90) {
    print(paste(student_name, "'s average grade this semester is", round(average_score, 2)))
  }
}
```
#c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests.
```{r}
for (j in 2:ncol(grades)) {  # Start from column 2 (Grade1) to avoid the Name column
  test_average <- mean(grades[, j])
  if (test_average < 80) {
    print(paste("The test", j - 1, " was difficult."))
  }
}
```
#d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.
```{r}
for (i in 1:nrow(grades)) {
  student_name <- grades$Name[i]
  highest_score <- max(grades[i, 2:5])
  if (highest_score > 90) {
    print(paste(student_name, "'s highest grade this semester is", highest_score))
  }
}
```

