---
title: "Worksheet 3b in R"
author: "Erl Syron R. Espadon"
output: pdf_document
---

#1. Create a data frame from the table
#a. codes
```{r}
df <- data.frame(
  Respondents = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20),
  Sex = c(2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2),
  Fathers_Occupation = c(1, 3, 3, 3, 1, 2, 3, 1, 1, 1, 3, 2, 1, 3, 3, 1, 3, 1, 2, 1),
  Persons_at_Home = c(5, 7, 3, 8, 5, 9, 6, 7, 8, 4, 7, 5, 4, 7, 8, 8, 3, 11, 7, 6),
  Siblings_at_school = c(6, 4, 4, 1, 2, 1, 5, 3, 1, 2, 3, 2, 5, 5, 2, 1, 2, 5, 3, 2),
  Types_of_houses = c(1, 2, 3, 1, 1, 3, 3, 1, 2, 3, 2, 3, 2, 2, 3, 3, 3, 3, 3, 2)
)
```
#b. Describe the data
```{r}
str(df)
summary(df)
```
#c. Calculate the mean number of siblings attending school
```{r}
mean(df$Siblings_at_school)
```
#d. Extract the first two rows and all columns
```{r}

df[1:2, ]
```
#e. Extract the 3rd and 5th row with the 2nd and 4th column
```{r}

df[c(3, 5), c(2, 4)]
```
#f. Select the variable 'Types_of_houses' and store it in a vector called 'types_houses'
```{r}
types_houses <- df$Types_of_houses
```
#g. Select only all Males respondent that their father occupation was farmer.
```{r}
# Filter the data frame to select only male respondents (Sex == 1) 
# whose fathers are farmers (Fathers_Occupation == 1)
males_farmer <- df[df$Sex == 1 & df$Fathers_Occupation == 1, ]

# Print the filtered data frame
print(males_farmer)
```
#h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. 
```{r}
# Filter the data frame to select only female respondents (Sex == 2)
# who have 5 or more siblings attending school (Siblings_at_school >= 5)
females_siblings <- df[df$Sex == 2 & df$Siblings_at_school >= 5, ]

# Print the filtered data frame
print(females_siblings)
```



#2. Write a R program to create an empty data frame. Using the following codes:
```{r}
df = data.frame(Ints=integer(),
Doubles=double(),Characters=character(),
Logicals=logical(), Factors=factor(), stringsAsFactors=FALSE)
print("Structure of the empty dataframe:") 
print(str(df))

#the code creates an empty data frame with columns pre-defined for different data types, ready to be filled with data later.
```


#3
# Create a data frame from the table
```{r}
df <- data.frame(
  Respondents = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
  Sex = c("Male", "Female", "Female", "Male", "Male", "Female", "Female", "Male", "Female", "Male"),
  Fathers_Occupation = c("Farmer", "Driver", "Others", "Others", "Farmer", "Driver", "Driver", "Others", "Farmer", "Others"),
  Persons_at_Home = c(1, 2, 3, 3, 1, 2, 2, 3, 1, 3),
  Siblings_at_School = c(5, 7, 3, 8, 6, 4, 4, 2, 11, 6),
  Types_of_Houses = c("Wood", "Congrete", "Congrete", "Wood", "Semi-congrete", "Semi-congrete", "Wood", "Semi-congrete", "Semi-congrete", "Congrete")
)
```
# 3. Create a .csv file of this. Save it as HouseholdData.csv
```{r}
write.csv(df, "HouseholdData.csv", row.names = FALSE)
```
#a. Import the csv file into the R environment. Write the codes.
```{r}
df <- read.csv("HouseholdData.csv")
```
# b. Convert the Sex into factor using factor() function and change it into integer.[Legend: Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
df$Sex <- factor(df$Sex, levels = c("Male", "Female"), labels = c(1, 2))
print(df$Sex)
```
# c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood = 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r}
df$Types_of_Houses <- factor(df$Types_of_Houses, levels = c("Wood", "Congrete", "Semi-congrete"), labels = c(1, 2, 3))
print(df$Types_of_Houses)
```
# d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What is the R code and its output?
```{r}
df$Fathers_Occupation <- factor(df$Fathers_Occupation, levels = c("Farmer", "Driver", "Others"), labels = c(1, 2, 3))
print(df$Fathers_Occupation)
```
# e. Select only all females respondent that has a father whose occupation is driver. Write the codes and its output.
```{r}
females_driver_father <- df[df$Sex == 2 & df$Fathers_Occupation == 2, ]
print(females_driver_father)
```
# f. Select the respondents that have greater than or equal to 5 number of siblings attending school. Write the codes and its output.
```{r}
respondents_5_siblings <- df[df$Siblings_at_School >= 5, ]
print(respondents_5_siblings)
```


#4. Interpret the graph
# Install the jpeg package if you haven't already
install.packages("jpeg")

```{r}
# Load the jpeg package
library(jpeg)

# Read the JPEG image
image <- readJPEG("C:\\Users\\Acer\\OneDrive\\Pictures\\Picturee.jpg") 

# Display the image
plot(as.raster(image))

# Example interpretation text
   interpretation <- "The graph displays the sentiment of tweets per day from July 14th to July 21st, 2020. Each day is represented by a group of three bars, representing the count of tweets with a negative, neutral, or positive sentiment.
Overall Trend: The graph shows a general trend of negative sentiment dominating throughout the week, with a slight increase in positive sentiment on July 20th.
Day-by-Day Breakdown:
July 14th: The majority of tweets were negative, followed by neutral and then positive.
July 15th: The number of negative tweets increased significantly, followed by neutral and then positive.
July 17th: The number of negative tweets remained high, followed by positive and then neutral.
July 18th: The number of negative tweets decreased slightly, followed by positive and then neutral.
July 20th: The number of negative tweets remained relatively low, followed by neutral and then positive.
July 21st: The number of negative tweets increased again, followed by positive and then neutral."

# Position the text (adjust coordinates as needed)
text(x = 1, y = 1, labels = interpretation, adj = c(.1, 0), cex = .5)
```