---
title: "Worksheet 4b in R"
author: "Erl Syron R. Espadon"
output: pdf_document
---

#Using Loop Function: for() loop
#1.	Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix. Hint: Use abs() function to get the absolute value
```{r}
# Create vectorA
vectorA <- c(1, 2, 3, 4, 5)

# Create a 5x5 zero matrix
matrix <- matrix(0, nrow = 5, ncol = 5)

# Use for loop to populate the matrix
for (i in 1:5) {
  for (j in 1:5) {
    matrix[i, j] <- abs(vectorA[i] - vectorA[j])
  }
}
print(matrix)
```


#2.	Print the string "*" using for() function. 
```{r}
# Loop for each row
for (i in 1:5) {
  for (j in 1:i) {
    cat("* ")
  }
  cat("\n")
}
```


#3.	Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r fibonacci}
# Define a function to generate the Fibonacci sequence
fibonacci_sequence <- function(start_num) {
  # Initialize variables for the Fibonacci sequence
  a <- 0
  b <- start_num

  # Use repeat and break statements to generate the sequence
  repeat {
    # Print the current number
    cat(a, " ")

    # Calculate the next number in the sequence
    c <- a + b

    # Update the variables for the next iteration
    a <- b
    b <- c

    # Break the loop if the current number exceeds 500
    if (c > 500) {
      break
    }
  }
}

# Get user input for the starting number
start_num <- as.numeric(readline("Enter the starting number for the Fibonacci sequence: "))


# Call the function to generate the sequence
fibonacci_sequence(start_num)
```


#4. Import the dataset as shown in Figure 1 you have created previously.
```{r}
View(df)
```

