---
title: "Worksheet 3a in R"
author: "Erl Syron R. Espadon"
output: pdf_document
---

```{r}
LETTERS
```

#1. VECTORS

#a. Produce a vector that contains the first 11 letters.
```{r}
first_11_letters <- LETTERS[1:11]
first_11_letters

```
#b. Produce a vector that contains the odd-numbered letters.
```{r}
odd_letters <- LETTERS[seq(1, length(LETTERS), by=2)]
odd_letters

```
#c. Produce a vector that contains the vowels.
```{r}
vowels <- LETTERS[c(1, 5, 9, 15, 21)]
vowels

```
#d. Produce a vector that contains the last 5 lowercase letters.
```{r}
last_5_lowercase <- letters[22:26]
last_5_lowercase

```
#e. Produce a vector that contains letters between 15 to 24 in lowercase.
```{r}
mid_lowercase <- letters[15:24]
mid_lowercase

```

#2. Create a vector with the average temperatures in April for Tuguegarao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City.
#a. Character vector for the cities.
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
city

```
#b. Temperature vector.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
temp

```
#c. Combine the city and temp vectors into a dataframe.
```{r}
weather_df <- data.frame(City = city, Temperature = temp)
weather_df

```
#d. Rename the columns using names() function.
```{r}
names(weather_df) <- c("City", "Temperature")
weather_df

```
#e. Print the structure using str() function.
```{r}
str(weather_df)

```
#f. Content of row 3 and row 4.
```{r}
row_3_4 <- weather_df[3:4, ]
row_3_4

```
#g. Display the city with the highest and lowest temperature.
```{r}
highest_temp <- weather_df[which.max(weather_df$Temperature), ]
lowest_temp <- weather_df[which.min(weather_df$Temperature), ]
highest_temp
lowest_temp

```

#MATRICES
#1. a. Create a matrix of numbers one to eight and eleven to fourteen with four columns and three rows.
```{r}
matrix1 <- matrix(c(1:8, 11:14), ncol = 4, byrow = TRUE)
matrix1

```
#b. Multiply the matrix by two.
```{r}
matrix_multiplied <- matrix1 * 2
matrix_multiplied

```
#wc. Content of row 2.
```{r}
row_2 <- matrix1[2, ]
row_2

```
#d. Display columns 3 and 4 in rows 1 and 2.
```{r}
cols_3_4_rows_1_2 <- matrix1[1:2, 3:4]
cols_3_4_rows_1_2

```
#e. Display only columns 2 and 3, row 3.
```{r}
cols_2_3_row_3 <- matrix1[3, 2:3]
cols_2_3_row_3

```
#f. Display only column 4.
```{r}
col_4 <- matrix1[, 4]
col_4

```
#g. Name the rows and columns.
```{r}
dimnames(matrix_multiplied) <- list(c("isa", "dalawa", "tatlo"), c("uno", "dos", "tres", "quatro"))
matrix_multiplied

```
#h. Reshape the matrix using dim().
```{r}
dim(matrix1) <- c(6, 2)
matrix1

```

#ARRAYS
#a.Create an array with the specified numeric values, repeated twice, and a three-dimensional array with 4 columns and 2 rows.
```{r}
values <- c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1)
array_data <- array(rep(values, each = 2), dim = c(2, 4, 3))
array_data

```
#b. Dimensions of the array.
```{r}
dim(array_data)

```
#c. Name the rows and columns.
```{r}
dimnames(array_data) <- list(letters[1:2], LETTERS[1:4], c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array"))
array_data

```

